From physical principles to ???

### 2016-04-29 2015CompleteDerivation/completeDerivation.tex
Another pass at equations of motion

	The first problem is methodological. To be able to talk about the states of a part we need a physical process that is able to distinguish between them. For a pool ball we can imagine marking one spot with a different color. This allows us to track it as the ball moves or collides with other balls. For an electron, instead, we do not have a way to mark a piece. In fact when two electrons scatter we can't even tell which is which, let alone what part went where. The classical assumption may not hold because we do not have suitable physical processes at our disposal.

@@

We are finally ready to write the equations of motions. As per \ref{detrevmap} our evolution is at least a self-homeomorphism $f:\mathcal{S} \leftrightarrow \mathcal{S}$. But this is not enough.

The evolution must map the distribution point-wise so that the value associated at the initial state is the same as the value. All of the system that starts in $\mathcal{s}$ has to end up in $\hat{\mathcal{s}}$. In math terms $\rho_{\hat{\mathcal{c}}} (\hat{\mathcal{s}}) = \rho_\mathcal{c}(\mathcal{s})$.

In the same way, if $U \in \mathcal{S}$ is a set of initial particle state and $\Lambda_U(\mathcal{c})$ is the amount of material associated to that set, then we expect it to be equal to the amount of material $\Lambda_{\hat{U}}(\hat{\mathcal{c}})$ associated with the set of final states $\hat{U} \in \mathcal{S}$.

But probably the best way to look at it is that initial and final set of states have to contain the same cardinality of states and possibilities, which where defined by our metric $\omega$. Therefore the area in each degree of freedom, which represents the number of possibilities in said d.o.f., needs to be mapped to an equal area in the transported d.o.f. (i.e. the d.o.f. defined by the transported state variables). Independent d.o.f. must remain independent, and therefore transported orthogonal d.o.f. remain orthogonal. This means that the product of possibilities of independent d.o.f. is also conserved. These statements are the physical justification of the Gromov's non-squeezing theorem and Liouville's theorem. And they give intuitive insight on the geometry of Hamiltonian systems.

Mathematically, under the classical Hamilotnian assumption,  deterministic and reversible evolution is an self-isometry. That is, it does not just preserve the topology but also the metric.

\begin{defn}\label{canonical_transformation}
	A deterministic and reversible evolution map for a homogeneous classical material is a canonical transformation and self-isometry, $\mathcal{T}_{\Delta t}:(T^*\mathcal{Q}, \omega) \rightarrow (T^*\mathcal{Q}, \omega)$.
\end{defn}

\begin{justification}
	We claim $\mathcal{T}_{\Delta t}$ is a self-homeomorphism. $\mathcal{T}_{\Delta t}$ is a deterministic and reversible evolution map and \ref{detrevmap}.
	We claim $\mathcal{T}_{\Delta t}$ is an self-isometry on $(T^*\mathcal{Q}, \omega)$. The Jacobian for $\mathcal{T}_{\Delta t}$ has to exist and be different from zero in order for the distribution on final states to be defined. $\mathcal{T}_{\Delta t}$ is a diffeomorphism.  $\omega$ is the metric for the cardinality of possibilities. A deterministic and reversible process conserves the number of states and possibilities. $\omega$ is invariant under a deterministic and reversible evolution map.
\end{justification}

@@

>>>>

	The first problem is methodological. To be able to talk about the states of a part we need a physical process that is able to distinguish between them. For a pool ball we can imagine marking one spot with a different color. This allows us to track it as the ball moves or collides with other balls. For an electron, instead, we do not have a way to mark a piece. In fact when two electrons scatter we can't even tell which is which, let alone what portion went where. The classical assumption may not hold because we do not have suitable physical processes at our disposal.

@@

We are finally ready to write the equations of motion. As per \ref{detrevmap} our evolution is at least a self-homeomorphism $f:\mathcal{S} \leftrightarrow \mathcal{S}$. But this is not enough.

The evolution must map the distribution point-wise so that the value associated at the initial state is the same as the value at the final state. All of the system that starts in $\mathcal{s}$ has to end up in $\hat{\mathcal{s}}$. In math terms $\rho_{\hat{\mathcal{c}}} (\hat{\mathcal{s}}) = \rho_\mathcal{c}(\mathcal{s})$.

In the same way we expect the total amount of material to be conserved. If $U \in \mathcal{S}$ is a set of initial particle states and $\Lambda_U(\mathcal{c})$ is the amount of material associated with that set, then we expect it to be equal to the amount of material $\Lambda_{\hat{U}}(\hat{\mathcal{c}})$ associated with the set of final states $\hat{U} \in \mathcal{S}$.

But probably the best way to look at it is that initial and final sets of states have to possess the same cardinality of states and possibilities, which were defined by our metric $\omega$. Therefore the area within a degree of freedom, which represents the number of possibilities in said d.o.f., needs to be mapped to an equal area within the transported d.o.f. (i.e. the d.o.f. defined by the transported state variables). Independent d.o.f. must remain independent, and therefore transported orthogonal d.o.f. remain orthogonal. This means that the product of possibilities of independent d.o.f. is also conserved. These statements are the physical justification of Gromov's non-squeezing theorem and Liouville's theorem. And they give intuitive insight on the geometry of Hamiltonian systems.

Mathematically, under the classical Hamiltonian assumption,  deterministic and reversible evolution is a self-symplectomorphism (or self-isometry or canonical transformation depending on your math training). That is, it does not just preserve the topology but also the symplectic metric $\omega$.

\begin{defn}\label{canonical_transformation}
	A deterministic and reversible evolution map for a homogeneous classical material is self-symplectomorphism. That is: $\mathcal{T}_{\Delta t}: T^*\mathcal{Q} \rightarrow T^*\mathcal{Q}$ and $\mathcal{T}_{\Delta t}^*\omega = \omega$ where $\mathcal{T}_{\Delta t}^*$ is the pullback of $\mathcal{T}_{\Delta t}$.
\end{defn}

\begin{justification}
	We claim $\mathcal{T}_{\Delta t}$ is a self-homeomorphism on $T^*\mathcal{Q}$. The state space for particles of a homogeneous classical material is $T^*\mathcal{Q}$ by \ref{classical_phase_space} $\mathcal{T}_{\Delta t}$ is a deterministic and reversible evolution map and \ref{detrevmap}.

	We claim $\mathcal{T}_{\Delta t}$ is a symplectomorphism on $(T^*\mathcal{Q}, \omega)$. The distribution on final states must be defined. The Jacobian for $\mathcal{T}_{\Delta t}$ exists and is non-zero. $\mathcal{T}_{\Delta t}$ is a diffeomorphism. A deterministic and reversible process conserves the number of states and possibilities. $\omega$ is the metric for the cardinality of possibilities. $\omega$ is invariant under a deterministic and reversible evolution map. $\mathcal{T}_{\Delta t}$ is a symplectomorphism by definition.
\end{justification}

@@

It's important to realize that, during the derivation, multiple mathematical features (e.g. invariant densities, cotangent bundle for phase space, symplectomorphism) where justified by the same physical assumption. As such, the math itself does not give us a conceptually unified picture, it gets lost. This is one of the reason we believe that starting from the physical description is objectively better if we are to come to a better understanding of our physical theories.

Note that we could have taken different approaches. For example, we could have appealed to information theory and asked our invariant distributions to preserve Shannon's information entropy, as no information should be gained or lost during a deterministic and reversible process. Or we could have appealed to statistical mechanics and required that the determinant of the covariance matrix be conserved, as deterministic and reversible process should be defined at the same level of precision. With suitable treatment of independent d.o.f. both these approaches recover Hamiltonian mechanics as well. While the full treatment goes outside the scope of this work, we want to underline that the definitions used in this work go a long way to building bridges at the core of different mathematical and scientific disciplines.
